---
title: "W04-summative"
output: html
self-contained: true
---

# W04 summative:

#⚙️Setup

```{r setup, message=FALSE, warning=FALSE}
library(mosaic)
library(dplyr)
library(readr)
library(lubridate)
library(tidymodels)
```

#📂opening the dataset

```{r, message=FALSE, warning=FALSE }
filepath <- "C:/Users/hanam/Desktop/DS202A/data/UK-HPI-full-file-2023-06.csv"
uk_hpi <- read_csv(filepath)
```

#🗃️pre-processing data (step 1)

```{r}
selected_cols    <- c("Date", "RegionName", "12m%Change", "1m%Change", "SalesVolume")

df_uk <-
    uk_hpi %>%
    select(all_of(selected_cols)) %>%
    rename(date = Date, region = RegionName, yearly_change = `12m%Change`, monthly_change = `1m%Change`) %>%
  filter(region == "United Kingdom") %>%
    drop_na() %>% 
    mutate(date = lubridate::dmy(date))

df_uk  %>% head(8) %>% knitr::kable()
```

#adding lagged variables

```{r}
df_uk <- 
    df_uk %>%
    arrange(date) %>% 
    mutate(SalesVolume1 = lag(SalesVolume, 1)) %>%
   mutate(monthly_changes_lag1 = lag(monthly_change, 1)) %>%
    drop_na() %>% 
    arrange(desc(date))

```

#separating data- testing/training data sets (step 2)

```{r, message=FALSE, warning=FALSE}
df_uk_train <- df_uk %>% filter(date <   ymd("2020-01-01"))
df_uk_test  <- df_uk %>% filter(date >=  ymd("2020-01-01"))
```

Here I am creating a training set (df_uk_train) which contains all data for the UK up until December 2019, and a testing set (df_uk_test) contain data from January 2020 onwards.

# linear regression- Sales volume increase per month

#Building linear model- model 1 (Step 3 & 4)

Here I am building a linear model in which 'SalesVolume' is our outcome variable and 'SalesVolume1' (lagged sales volume) is our predictor, in order to measure change in sales volume per 1 month.

```{r}
model1 <- 
  linear_reg() %>%
    set_engine("lm") %>%
    fit(SalesVolume ~ SalesVolume1, data = df_uk_train)
```

```{r}
##model coefficients
model1 %>% tidy()
```

The value of our intercept is approximately 1.36, whereas our gradient is 8.37. Therefore we can predict for every 1 month, sales volume will increase by 8.37%.

If the model is a good fit, we can expect the slope to be correct within a 95% confidence interval of 8.37 +/- 4.09.

```{r}
##full model summary
model1$fit %>% summary()
```

```{r}
#Overall statistics
model1 %>% glance()
```

#Building a linear model- model2

Here I am building a linear model with the variables 'SalesVolume' as the outcome variable and 'monthly_change' as the predictor variable.

```{r}
model2 <- 
  linear_reg() %>%
    set_engine("lm") %>%
    fit(SalesVolume ~ monthly_changes_lag1, data = df_uk_train)
```

```{r}
##model coefficients
model2 %>% tidy()

```

In model 2, our intercept is 79985, and our gradient is 13020. Therefore we can predict for every month, sales volume will increase by 13020 units.

```{r}
##full model summary
model2$fit %>% summary()
```

```{r}
#Overall statistics
model2 %>% glance()
```

## 📈Plots/diagrams

#Training set-model 1 (step 5)

```{r}
plot_df <- model1 %>% augment(df_uk_train)
g <- ggplot(plot_df, aes(x = .pred, y = .resid)) +
    geom_point(alpha=0.2, size=3, color="red", stroke=1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(x = "Fitted values", y = "Residuals", title="Residuals vs Fitted") + 
    theme_bw() + 
    theme(axis.title.x = element_text(size = rel(1.2)), 
          axis.text.x = element_text(size = rel(1.2)),
          axis.title.y = element_text(size = rel(1.2)),
          axis.text.y = element_text(size = rel(1.2)),
          plot.title = element_text(size = rel(1.5), face = "bold"))
g
```

The model appears to be a good fit, as the residual plot is linear.

#Testing set- model 1

```{r}
ggplot(model1 %>% augment(df_uk_test), aes(.pred, .resid)) + 
    geom_point(alpha=0.2, size=3, color="red", stroke=1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(x = "Fitted values", y = "Residuals", title="Residuals vs Fitted") + 
    theme_bw() + 
    theme(axis.title.x = element_text(size = rel(1.2)), 
          axis.text.x = element_text(size = rel(1.2)),
          axis.title.y = element_text(size = rel(1.2)),
          axis.text.y = element_text(size = rel(1.2)),
          plot.title = element_text(size = rel(1.5), face = "bold"))

```

The model also appears to be a good fit for the testing data frame. It appears to be linear, however with a slight curvature and a larger scattering than our training set.

#Training set- model 2

```{r}
ggplot(model2 %>% augment(df_uk_train), aes(.pred, .resid)) + 
    geom_point(alpha=0.2, size=3, color="red", stroke=1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(x = "Fitted values", y = "Residuals", title="Residuals vs Fitted") + 
    theme_bw() + 
    theme(axis.title.x = element_text(size = rel(1.2)), 
          axis.text.x = element_text(size = rel(1.2)),
          axis.title.y = element_text(size = rel(1.2)),
          axis.text.y = element_text(size = rel(1.2)),
          plot.title = element_text(size = rel(1.5), face = "bold"))
```

In comparison to our 'SalesVolume/SalesVolume1' lagged model, model 2 is less linear and highly scattered.

#Testing set- model 2

```{r}
ggplot(model2 %>% augment(df_uk_test), aes(.pred, .resid)) + 
    geom_point(alpha=0.2, size=3, color="red", stroke=1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(x = "Fitted values", y = "Residuals", title="Residuals vs Fitted") + 
    theme_bw() + 
    theme(axis.title.x = element_text(size = rel(1.2)), 
          axis.text.x = element_text(size = rel(1.2)),
          axis.title.y = element_text(size = rel(1.2)),
          axis.text.y = element_text(size = rel(1.2)),
          plot.title = element_text(size = rel(1.5), face = "bold"))
```

There doesn't appear to be much of a linear relationship between our variables in model 2 testing set.

# Evaluating model:

#🖩Calculating MAE- Training set/model 1

The mean absolute error allows us to distinguish how far the model is off from a true observation.

```{r}
model1 %>% 
    augment(df_uk_train) %>%
    mae(truth = SalesVolume, estimate = .pred)
```

```{r}
favstats(~SalesVolume, data = df_uk)
```

As demonstrated by the summary descriptives of sales volume provided by favstats(), we can see the central tendency is 27,648. The MAE in comparison is only a fraction of this value at 9228, indicating a relatively accurate prediction with an average error only approximately 9000 away from the mean.

#Calculating MAE- Testing set/model 1 (Step 6)

```{r}
model1%>% 
    augment(df_uk_test) %>%
    mae(truth = SalesVolume, estimate = .pred)
```

The MAE for the testing set is significantly larger than the training set, indicating that the model is not generalizing well to the rest of our dataset.

#Calculating MAE- training set/model 2

```{r}
model2%>% 
    augment(df_uk_train) %>%
    mae(truth = SalesVolume, estimate = .pred)
```

Our second linear model has a larger MAE than model 1, additionally the training set doesn't seem to fit the linear model well.

#Calculating MAE- testing set/model 2

```{r}
model2%>% 
    augment(df_uk_test) %>%
    mae(truth = SalesVolume, estimate = .pred)
```

# Commentary (step 7)

I ran a linear model with the outcome variable as 'SalesVolume' and the explanatory variable as 'SalesVolume1' (i.e., the lagged sales volume). This made up my first model. By creating a linear regression with these variables we can analyze the monthly lag/ change in sales volume per 1 month. Running a regression between a variable and their lagged counterpart is necessary for analyze time-lapsed relationships hence why this was my first model. The MAE was relatively small for our training model but not testing set, indicating the linear model is potentially overfitted.

Other configurations I tried included a linear model for 'lm(SalesVolume \~ monthly_changes_lag1)'. This curated my model2 as the MAE for my first model was quite high and I just thought to attempt other potential logical configurations to analyze. However, the MAE was larger than model 1 for both the training and testing set, and the plot actually appeared to be less linear for both the training and testing data frame in our first model.

From this I can conclude model 1 was closer to a true model in analyzing sales volume per month.
