---
title: "path analysis"
output: html_notebook
---

```{r}
install.packages("psych")
install.packages("naniar")
install.packages("sjlabelled")
install.packages("tidyverse")
install.packages("mice")
install.packages("VIM")
install.packages("https://cran.r-project.org/src/contrib/Archive/MissMech/MissMech_1.0.2.tar.gz", repos=NULL, type="source")
install.packages("mice")
install.packages("MoEClust")
install.packages("miceadds")
install.packages("Rcpp")
```

```{r}
library(psych)
library(naniar)
library(sjlabelled)
library(tidyverse)
library(mice)
library(VIM)
library(MissMech)
library(mice)
library(MoEClust)
library(miceadds)
```

```{r}
college <- read_csv("C:/Users/h3man/OneDrive/Desktop/pb230/PB230-master/Secondary Data Analysis/Task 1/college.csv")
```

```{r}
describe(college)
```

```{r}
college <-
  college %>%
  replace_with_na(replace = list(open2 = 8:55, swb1 = 8:77)) 
describe(college)
```

```{r}
college_var <-
  college %>%
  select(4:14) 
md.pattern(college_var)
```

```{r}
aggr(college_var, col=c('white','red'), numbers=TRUE, sortVars=TRUE, cex.axis=.7, gap=3, ylab=c("Percentage of missing data","Missing Data Pattern"))
vis_miss(college_var)
```

```{r}
out.MCAR.ws <- TestMCARNormality(college_var, del.lesscases = 1)
summary(out.MCAR.ws)
```

data is missing completely at random 157 complete cases. 211 in total.

```{r}
college_var$na_count <- apply(is.na(college_var), 1, sum)
college_var
```

```{r}
college_var <- 
college_var %>% 
filter(na_count <= "1")
college_var
```

##scale mean imputation

```{r}
college_var <-
  college_var %>%
  rowwise()%>%
  mutate(meanopen = mean(c(open1,open2,open3,open4), na.rm = TRUE)) %>%
  mutate(meansc = mean(c(sc1,sc2,sc3,sc4), na.rm = TRUE)) %>%
  mutate(meanswb = mean(c(swb1,swb2,swb3), na.rm = TRUE)) 
college_var
```

```{r}
##open 

college_var <- within(college_var, open1 <- ifelse(is.na(open1), meanopen, open1))
college_var <- within(college_var, open2 <- ifelse(is.na(open2), meanopen, open2))
college_var <- within(college_var, open3 <- ifelse(is.na(open3), meanopen, open3))
college_var <- within(college_var, open4 <- ifelse(is.na(open4), meanopen, open4))

##sc

college_var <- within(college_var, sc1 <- ifelse(is.na(sc1), meansc, sc1))
college_var <- within(college_var, sc2 <- ifelse(is.na(sc2), meansc, sc2))
college_var <- within(college_var, sc3 <- ifelse(is.na(sc3), meansc, sc3))
college_var <- within(college_var, sc4 <- ifelse(is.na(sc4), meansc, sc4))

##swb

college_var <- within(college_var, swb1 <- ifelse(is.na(swb1), meanswb, swb1))
college_var <- within(college_var, swb2 <- ifelse(is.na(swb2), meanswb, swb2))
college_var <- within(college_var, swb3 <- ifelse(is.na(swb3), meanswb, swb3))

college_var
```

```{r}
md.pattern(college_var)
```

```{r}
college_var <-
  college_var %>%
  rowwise()%>%
  mutate(meanopen = mean(c(open1,open2,open3,open4))) %>%
  mutate(meansc = mean(c(sc1,sc2,sc3,sc4))) %>%
  mutate(meanswb = mean(c(swb1,swb2,swb3)))
college_var
```

##univariate outliers

```{r}
# Standardize variables

college_var$zopen <- scale(college_var$meanopen)
college_var$zsc <- scale(college_var$meansc)
college_var$zswb <- scale(college_var$meanswb)

# Remove open outliers

college_var <- 
  college_var %>%
  filter(zopen >= -3.30 & zopen <= 3.30)
college_var

# Remove sc outliers

college_var <- 
  college_var %>%
  filter(zsc >= -3.30 & zsc <= 3.30)
college_var

# Remove swb outliers

college_var <- 
  college_var %>%
  filter(zswb >= -3.30 & zswb <= 3.30)
college_var
```

##multivariate outliers

```{r}
linear.model <- lm(open1 ~ meanopen + meansc + meanswb, data=college_var)
college_var$res  <- college_var$open1 - predict(linear.model) 
college_var$mahal <- MoE_mahala(linear.model, college_var$res)
college_var 
summary(linear.model)
```

```{r}
# Remove multivariate outliers

college_var <- 
  college_var %>%
  filter(mahal <= 16.27)
college_var
```

```{r}
describe(college_var$meanopen)
describe(college_var$meansc)
describe(college_var$meanswb)
```

openness had a mean of 5.34. openness had a slight negative skew (-0.68) and a negative kurtosis (-0.1). self-compassion has a mean of 5.57 slightly higher than both other variables. a symmetrical skew of -0.43, and a slight negative kurtosis of -0.33. subjective well-being had a mean of 5.4, a symmetrical skew of -0.18, and a slight negative kurtosis of -0.48.

##distributions

```{r}
library(ggplot2)
ggplot(college_var, aes(x=meanopen)) +
   geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
   geom_vline(aes(xintercept=mean(meanopen)),
            color="blue", linetype="dashed", size=1) +
   labs(title="openness distribution",x="openness", y = "density")+
  theme_classic()
```

```{r}
ggplot(college_var, aes(x=meansc)) +
   geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  geom_vline(aes(xintercept=mean(meansc)),
            color="blue", linetype="dashed", size=1) +
   labs(title="self-compassion distribution",x="self-compassion", y = "density")+
  theme_classic()
```

```{r}
ggplot(college_var, aes(x=meanswb)) +
   geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  geom_vline(aes(xintercept=mean(meanswb)),
            color="blue", linetype="dashed", size=1) +
   labs(title="subjective well-being distribution",x="subjective well-being", y = "density")+
  theme_classic()
```

*research question* is the relationship between openness and subjective well-being mediated by self-compassion *null hypothesis* the indirect effect of the relationship between openness and subjective well-being through self-compassion will be zero

```{r}
install.packages("OpenMx")
install.packages("lavaan")
install.packages("semPlot")
install.packages("mediation")
install.packages("Hmisc")
install.packages("performance")
```

```{r}
library("OpenMx")
library("lavaan")
library("semPlot")
library("performance")
library("mediation")
library("Hmisc")
```

##correlations

```{r}
rcorr(as.matrix(college_var[,c("meanopen","meansc", "meanswb")], type="pearson"))
```

the variables are all positively correlated, openness and self compassion were substantially moderated at 0.61. self-compassion and subjective well-being were moderately correlated at 0.34, and openness and subjective well-being were slightly correlated at 0.24.

##building path models

```{r}
mediation.model <-
  "meanswb ~ meanopen + meansc
meansc ~ meanopen"
```

```{r}
mediation.fit <- sem(mediation.model, data=college_var)
```

##just identified model

```{r}
med.model <-
"meanswb ~ c * meanopen + b * meansc
meansc ~ a * meanopen
ab := a*b
total := ab+c"
```

##bootstrapping

```{r}
just.model <- sem(med.model, data = college_var, se = "bootstrap", bootstrap = 5000)
```

```{r}
summary(just.model, fit.measures=TRUE, standardized=TRUE, ci=TRUE)
```

a just identified model. RMSEA and RMR is 0, CFI and TLI is 1. fit indexes are perfect which will become useful once we omit the c\* path from the model.

```{r}
parameterestimates(just.model, boot.ci.type = "perc", standardized = TRUE)
```

from the summary table we can see openness is a significant predictor of self-compassion (b= 0.4, CI= 0.32, 0.47) and self-compassion is a significant predictor of subjective well-bein (b= 0.4, CI= 0.18, 0.6). This implies mediation is occurring. However, the direct path between openness and swb is quite small (b=0.05 CI= -0.07, 0.17). the indirect effect (b=0.15, CI= 0.06, 0.25) is significant therefore we can reject the null hypothesis, self-compassion mediates the relationship between openness and swb.

```{r}
##linear model path a
model.a <- lm(meansc ~ 1 + meanopen, data = college_var)

##linear model for path b and c
model.b <- lm(meanswb ~ 1 + meanopen + meansc, data = college_var)

##mediation using linear regression
mediation <- mediate(model.a, model.b, treat="meanopen", mediator="meansc", boot=TRUE, sims=5000, boot.ci.type="perc", conf.level= 0.95) 
summary(mediation)
```
```{r}
lavInspect(just.model, what = "rsquare")
```

```{r}
semPaths(just.model, "par",
             sizeMan = 15, sizeInt = 15, sizeLat = 15,
             edge.label.cex=1.5,
             fade=FALSE)
```

##over identified model principal of model fit, omitting path C to see whether it is needed or if path ab suffices.

```{r}
med.model2 <- "
# the b path
  meanswb ~ b * meansc
# the a path
  meansc ~ a * meanopen
# indirect effect
ab := a*b
"
```

##fitting model

```{r}
over.model <- sem(med.model2, data = college_var, se = "bootstrap", bootstrap = 5000)
```

```{r}
summary(over.model, fit.measures=TRUE, standardized=TRUE, ci=TRUE)
```

our fit indexes indicate the model is an approximate fit and path c is not needed. RMSEA= 0, SRMR= 0.01. CFI= 1, TLI= 1.01. TLI and CFI should be above .90 (difference from worst possible fit)

RMSEA and SRMR should be below .10 (difference from perfect fit) recall from our just-identified model, the c path was small and non-significant (b=0.05, p=0.4).due to parsimony we shall omit path c.

the over-identified model fits the data well - the model implied correlation matrix well approximates the actual correlation matrix.

```{r}
## covariance matrix for the data
cov(as.matrix(college_var[,c("meanswb","meansc", "meanopen")]))

## model implied covariance matrix
fitted(over.model)
```

```{r}
parameterestimates(over.model, boot.ci.type = "perc", standardized = TRUE)
```

with 5000 resamples, the CI runs for our indirect effect of b=0.18 from 0.09 to 0.26.Therefore we can reject the null hypothesis, self-compassion mediates subjective well-being even when we leave out the direct effect.

```{r}
lavInspect(over.model, what = "rsquare")
```

r2 values, 11% of the variance in swb is explained by the path model. 36% of the variance in self compassion is explained by the path model.

```{r}
semPaths(over.model, "par",
             sizeMan = 15, sizeInt = 15, sizeLat = 15,
             edge.label.cex=1.5,
             fade=FALSE)
```

In order to test the hypothesized mediation model, path analysis with maximum likelihood estimation was conducted using the laavan R package. A fully mediated model was tested that included one independent variable (openness), one mediator variable (self-compassion), and one dependent variable (subjective well-being). Conventional cut-off criteria were adopted to indicate adequate fit for the hypothesised model to the observed data (TLI and CFI \> .90, RMSEA and SRMR \< .10).

The path model displayed adequate fit to the data (CFI = 1, TLI = 1.01, RMSEA = 0, SRMR = 0.01). The path coefficient between openness and self-compassion was significant (b = .40, B = .03, 95% bootstrap CI = .32, .47). Likewise, the path coefficient between self-compassion and subjective wellbeing was also significant (b = .44, B = .09, 95% bootstrap CI = .24, .62). To determine whether the mediated effect of openness on subjective well-being through self-compassion was statistically significant, the indirect effect (i.e., ab) and its 95% confidence interval was calculated using a bootstrapping procedure that drew 5,000 resamples. The indirect effect was ab = .18 and the bootstrapped confidence interval with 5,000 resamples did not include zero, 95% CI [.09,.26].
